---
title: 'Chopped ratings & ingredients'
subtitle: 'Tidy Tuesday 2020-08-25'
author: 'Shelby Bachman'
output: 
  html_document:
    theme: flatly
    highlight: default
---

```{r setup, include = FALSE}

rm(list = ls())

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggpubr)
library(extrafont)
library(patchwork)

```

Today's data comes from [Kaggle](https://www.kaggle.com/jeffreybraun/chopped-10-years-of-episode-data) courtesy of Jeffrey Braun and [Nick Wan](https://twitter.com/nickwan/status/1291586894145490944?s=20).

This week we are exploring data from **Chopped**, a reality-based cooking television game show series. In each episode of this show four chefs compete in a three-course contest, where they attempt to incorporate unusual combinations of ingredients into dishes that are evaluated by a panel of judges. At the beginning of each course (appetizer, entree, and dessert), the chefs are each given a basket containing four mystery ingredients and are expected to create dishes that use all of them in some way. 

Below I do some basic exploration of how ratings evolved across the lifetime of the show and the most frequent ingredients used in the show.

<hr>

```{r load_data, include = FALSE}

chopped <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-25/chopped.tsv')

```

```{r split_ingredients, include = FALSE}

# separate ingredients from each course
chopped <- chopped %>%
  rowwise() %>%
  separate(appetizer, into = c('app_1', 'app_2', 'app_3', 'app_4'), 
           sep = ', ', extra = 'merge') %>%
  separate(entree, into = c('entree_1', 'entree_2', 'entree_3', 'entree_4', 'entree_5'), 
           sep = ', ', extra = 'merge') %>%
  separate(dessert, into = c('dessert_1', 'dessert_2', 'dessert_3', 'dessert_4', 'dessert_5'),
           sep = ', ', extra = 'merge')

```

```{r fig_ratings, fig.width = 8}

ggplot(data = chopped %>% filter(season <= 40), 
       aes(x = season, y = episode_rating)) +
  geom_point(size = 2, alpha = 0.5, colour = '#9BC1BC') +
  geom_path(data = chopped %>% 
              filter(season <= 40) %>%
              group_by(season) %>% 
              summarise(episode_rating = mean(episode_rating, na.rm = TRUE)),
            size = 2) +
  labs(x = 'Season', y = 'Episode Rating') +
  scale_x_continuous(breaks = seq(5, 40, 5)) +
  scale_y_continuous(breaks = seq(5, 10, 1)) +
  theme_pubr() +
  theme(text = element_text(size=16,  family='Lato'),
        axis.text.x = element_text(size = 10, family='Lato'),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 10, family='Lato'),
        axis.ticks.y = element_blank()) +
  plot_annotation(
  title = "How did ratings of chopped evolve across all 40 seasons?",
  theme = theme(plot.title = element_text(
    size = 16, hjust = 0.5, family = 'Lato'))
  )

```
<hr>

```{r transform_data, include = FALSE}

chopped_long <- as.data.frame(
  rbind(
    chopped %>%
      select(season, season_episode, episode_rating, app_1:app_4) %>%
      pivot_longer(cols = app_1:app_4, 
                   names_to = 'ingredient_no', 
                   names_prefix = 'app_',
                   values_to = 'ingredient', 
                   values_drop_na = TRUE) %>%
      mutate(course = 'appetizer'),
    
    chopped %>%
      select(season, season_episode, episode_rating, entree_1:entree_5) %>%
      pivot_longer(cols = entree_1:entree_5, 
                   names_to = 'ingredient_no', 
                   names_prefix = 'entree_',
                   values_to = 'ingredient', 
                   values_drop_na = TRUE) %>%
      mutate(course = 'entree'),
    
    chopped %>%
      select(season, season_episode, episode_rating, dessert_1:dessert_5) %>%
      pivot_longer(cols = dessert_1:dessert_5, 
                   names_to = 'ingredient_no', 
                   names_prefix = 'dessert_',
                   values_to = 'ingredient', 
                   values_drop_na = TRUE) %>%
      mutate(course = 'dessert')
  )
) %>%
  mutate(ingredient = str_to_lower(ingredient))

summary_ingredients <- chopped_long %>%
  group_by(ingredient, course) %>%
  summarise(count = n())

summary_ingredients_all <- chopped_long %>%
  group_by(ingredient) %>%
  summarise(count = n()) 

```

```{r fig_ingredients, include = FALSE}

p_app <- ggplot(summary_ingredients %>% 
         filter(course == 'appetizer', count > 4),
       aes(x = reorder(ingredient, count), y = count)) +
  geom_segment(aes(xend = ingredient, yend = 4)) +
  geom_point(size = 4, colour = '#ED6A5A') +
  coord_cartesian(ylim = c(4, 10)) +
  scale_y_continuous(breaks = seq(4, 10)) +
  labs(x = '', y = 'frequency') +
  annotate('text', x = 7, y = 7.5, label = 'appetizers', 
           colour = '#ED6A5A', size = 6, family = 'Lato') +
  coord_flip() +
  theme_pubr() + 
  theme(text = element_text(size=10,  family='Lato'),
        axis.text.x = element_text(size = 10, family='Lato'),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 10, family='Lato'),
        axis.ticks.y = element_blank())

p_entree <- ggplot(summary_ingredients %>% 
         filter(course == 'entree', count > 4),
       aes(x = reorder(ingredient, count), y = count)) +
  geom_segment(aes(xend = ingredient, yend = 4)) +
  geom_point(size = 4, colour = '#E5DF61') +
  coord_cartesian(ylim = c(4, 12)) +
  scale_y_continuous(breaks = seq(4, 12)) +
  labs(x = '', y = 'frequency') +
  annotate('text', x = 11, y = 9, label = 'entrees', 
           colour = '#E5DF61', size = 6, family = 'Lato') +
  coord_flip() +
  theme_pubr() + 
  theme(text = element_text(size=10,  family='Lato'),
        axis.text.x = element_text(size = 10, family='Lato'),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 10, family='Lato'),
        axis.ticks.y = element_blank())

p_dessert <- ggplot(summary_ingredients %>% 
         filter(course == 'dessert', count > 4),
       aes(x = reorder(ingredient, count), y = count)) +
  geom_segment(aes(xend = ingredient, yend = 4)) +
  geom_point(size = 4, colour = '#5D576B') +
  coord_cartesian(ylim = c(4, 13)) +
  scale_y_continuous(breaks = seq(4, 13)) +
  labs(x = '', y = 'frequency') +
  annotate('text', x = 10, y = 9, label = 'dessert', 
           colour = '#5D576B', size = 6, family = 'Lato') +
  coord_flip() +
  theme_pubr() + 
  theme(text = element_text(size=10,  family='Lato'),
        axis.text.x = element_text(size = 8, family='Lato'),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 10, family='Lato'),
        axis.ticks.y = element_blank())


```

```{r fig_overall, include = FALSE}

p_all <- ggplot(summary_ingredients_all %>% 
         filter(count > 9),
       aes(x = reorder(ingredient, count), y = count)) +
  geom_segment(aes(xend = ingredient, yend = 9)) +
  geom_point(size = 4, colour = '#9BC1BC') +
  coord_cartesian(ylim = c(9, 15)) +
  scale_y_continuous(breaks = seq(9, 15)) +
  labs(x = '', y = 'frequency') +
  annotate('text', x = 13, y = 14, label = 'all courses', 
           colour = '#9BC1BC', size = 8, family = 'Lato') +
  coord_flip() +
  theme_pubr() + 
  theme(text = element_text(size=10,  family='Lato'),
        axis.text.x = element_text(size = 10, family='Lato'),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 10, family='Lato'),
        axis.ticks.y = element_blank())

```

```{r fig_compile, fig.width = 10, fig.height = 12}

layout <- p_all / (p_app | p_entree | p_dessert)

layout +
  plot_annotation(
  title = "What were the most commmon ingredients?",
  theme = theme(plot.title = element_text(
    size = 16, hjust = 0.5, family = 'Lato'))
  )

```