---
title: 'Threats & responses to plant extinction worldwide'
subtitle: 'Tidy Tuesday 2020-08-18'
author: 'Shelby Bachman'
output: 
  html_document:
    theme: sandstone
    highlight: tango
---

```{r setup, include = FALSE}

rm(list = ls())

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(janitor)
library(kableExtra)
library(ggpubr)
library(ggthemr)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
ggthemr('light')

```

Data used here comes from the [International Union for Conservation of Nature (IUCN) Red List of Threatened Species (Version 2020-1)](https://www.iucnredlist.org/). <br>It was prepared by [Florent Lavergne](https://www.behance.net/florentlavergne).

```{r load_data, include = FALSE}

data <- tidytuesdayR::tt_load('2020-08-18')
plants <- data$plants
rm(data)

```

```{r clean_data, include = FALSE}

# convert action data from wide to long & rename action types
actions <- plants %>%
  select(-contains('threat')) %>%
  pivot_longer(cols = contains('action'),
               names_to = 'action_type',
               names_prefix = 'action_',
               values_to = 'action_taken', 
               values_drop_na = TRUE) %>%
  mutate(action_type = case_when(
    action_type == 'LWP' ~ 'Land & Water Protection',
    action_type == 'SM' ~ 'Species Management',
    action_type == 'LP' ~ 'Law & Policy',
    action_type == 'RM' ~ 'Research & Monitoring',
    action_type == 'EA' ~ 'Education & Awareness',
    action_type == 'NA' ~ 'Current action unknown'
  ))

# conver threat data from wide to long and rename threat types
threats <- plants %>%
  select(-contains('action')) %>%
  pivot_longer(cols = contains('threat'),
               names_to = 'threat_type',
               names_prefix = 'threat_',
               values_to = 'threatened',
               values_drop_na = TRUE) %>%
  mutate(threat_type = case_when(
    threat_type == 'AA' ~ 'Agriculture & Aquaculture',
    threat_type == 'BRU' ~ 'Biological Resource Use',
    threat_type == 'RCD' ~ 'Commercial Development',
    threat_type == 'ISGD' ~ 'Invasive Species',
    threat_type == 'EPM' ~ 'Energy Production & Mining',
    threat_type == 'CC' ~ 'Climate Change',
    threat_type == 'HID' ~ 'Human Intrusions',
    threat_type == 'P' ~ 'Pollution',
    threat_type == 'TS' ~ 'Transportation Corridor',
    threat_type == 'NSM' ~ 'Natural System Modifications',
    threat_type == 'GE' ~ 'Geological Events',
    threat_type == 'NA' ~ 'Threat unknown'
  ))

```

```{r filter_data, include = FALSE}

# filter action data to include only plants for which action was taken
actions <- actions %>%
  filter(action_taken == 1)

# filter threat data to include only threatened plants
threats <- threats %>%
  filter(threatened == 1)

```

## What factors threaten plant existence?

```{r viz_threatened, echo = FALSE, fig.width = 5}

summary_threats <- threats %>%
  group_by(threat_type) %>%
  summarise(tot_threatened = sum(threatened, na.rm = TRUE), .groups = NULL)

ggplot(data = summary_threats, 
       mapping = aes(x = tot_threatened,
                     size = tot_threatened,
                     y = threat_type)) +
 #                    y = reorder(threat_type, tot_threatened))) +
  geom_point(aes(size = tot_threatened), col = '#8ABD91') +
  labs(x = 'Number of threatened plant species worldwide', y = '', size = '') +
  guides(size = FALSE)

```

### Threatened plants by country

The visualizations below includes only countries with more than 5 threatened species.

```{r viz_threatened_country, echo = FALSE}

summary_country_threat <- threats %>%
  group_by(country) %>%
  summarise(tot_threatened = sum(threatened, na.rm = TRUE), .groups = NULL) %>%
  left_join(threats %>% select(country, continent), by = 'country')

ggplot(data = summary_country_threat %>% filter(tot_threatened > 5), 
       mapping = aes(x = tot_threatened,
                     y = reorder(country, tot_threatened),
                     colour = factor(continent))) +
  labs(x = 'Number of threatened plant species', y = '', colour = 'continent') +
  geom_point(size = 3)

```
```{r map_threats, echo = FALSE}

world <- ne_countries(scale = "medium", returnclass = "sf")

world_joined <- world %>%
  left_join(summary_country_threat %>% filter(tot_threatened > 5) %>%
              select(name = country, tot_threatened), 
            by = 'name')

ggplot(data = world_joined) +
  geom_sf(aes(fill = tot_threatened)) +
  labs(fill = 'Threatened species')

```

## What actions are being taken to counter these threats?

```{r viz_actions, echo = FALSE, fig.width = 5, fig.height = 3}

summary_actions <- actions %>%
  group_by(action_type) %>%
  summarise(tot_actions = sum(action_taken, na.rm = TRUE), .groups = NULL)

ggplot(data = summary_actions, 
       mapping = aes(x = tot_actions,
                     size = tot_actions,
                     y = action_type)) +
  geom_point(aes(size = tot_actions), col = '#8ABD91') +
  labs(x = 'Number of actions taken worldwide', y = '', size = '') +
  guides(size = FALSE)

```

### Actions taken by country

The visualization below includes only countries which have taken more than 3 actions and does not include reports in which the action type was unknown.

```{r viz_action_country, echo = FALSE}

summary_country_action <- actions %>%
  filter(action_type == 'Current action unknown') %>%
  group_by(country) %>%
  summarise(tot_action = sum(action_taken, na.rm = TRUE), .groups = NULL) %>%
  left_join(actions %>% select(country, continent), by = 'country')

ggplot(data = summary_country_action %>% filter(tot_action > 3),
       mapping = aes(x = tot_action,
                     y = reorder(country, tot_action),
                     colour = factor(continent))) +
  labs(x = 'Number of actions taken', y = '', colour = 'continent') +
  geom_point(size = 3)

```

```{r map_actions, echo = FALSE}

world_joined <- world %>%
  left_join(summary_country_action %>% filter(tot_action > 3) %>%
              select(name = country, tot_action), 
            by = 'name')

ggplot(data = world_joined) +
  geom_sf(aes(fill = tot_action)) +
  labs(fill = 'Actions taken')

```