---
title: "Erasmus exchange program participation"
author: "Shelby Bachman"
subtitle: "Tidy Tuesday 2021-03-08"
output:
  html_document:
    theme: flatly
    highlight: default
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

rm(list = ls())
library(dplyr)
library(tidyr)
library(stringr)
library(ggforce)
library(ggplot2)

```
  
```{r load_data}

erasmus <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-08/erasmus.csv')

erasmus <- erasmus %>%
  select(project_reference, 
         mobility_start_month, mobility_end_month,
         sending_country_code, receiving_country_code) %>%
  rowwise() %>%
  mutate(start_year = as.numeric(str_split(mobility_start_month, '-')[[1]][1]),
         end_year = as.numeric(str_split(mobility_end_month, '-')[[1]][1]))

```

Today's data comes from [Data.Europa](https://data.europa.eu/data/datasets?locale=en&catalog=eac&query=erasmus&page=1&sort=issued%2Bdesc,%20relevance%2Bdesc,%20title.en%2Basc) courtsey of [Data Is Plural](https://www.data-is-plural.com/archive/2022-02-09-edition/) and includes metrics on participation in the Erasmus exchange program at EU universities from the years `r range(erasmus$start_year)`. I was particularly interested in exploring the most popular destinations of students from each country, so I used the `ggforce` package to examine sending countries (left) versus receiving countries (right) for Erasmus program participants. I restricted the data to cases where students visited a university outside their home country.

<hr>

Since there are a lot of countries, I first examined the <u>destination countries</u> for students from the 5 countries that sent the most students on exchange visits.

```{r fig_receiving, fig.height = 12, fig.width = 10}

# determine 5 countries with most participants
top_countries <- erasmus %>%
  # only include international exchanges
  filter(sending_country_code != receiving_country_code) %>%
  group_by(sending_country_code) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  slice_head(n=5)

# create a custom color palette based on the top 5 countries
countries_pal <- c('DE' = '#000000', 
                   'ES' = '#ffc400',
                   'IT' = '#009246', 
                   'PL' = '#dc143c',
                   'UK' = '#00247d')

# tally counts of destinations by origin country
# format data for ggforce
erasmus_tidier <- erasmus %>%
  # only include international exchanges
  filter(sending_country_code != receiving_country_code) %>%
  # only include top 5 countries as determined above
  filter(sending_country_code %in% top_countries$sending_country_code) %>%
  group_by(sending_country_code, receiving_country_code) %>%
  summarize(freq = n()) %>%
  ungroup() %>%
  gather_set_data(x = 1:2) %>%
  arrange(x, sending_country_code, receiving_country_code)
erasmus_tidier$x <- factor(erasmus_tidier$x,
                           levels = c('sending_country_code', 'receiving_country_code'))

# create parallel sets plot
# showing destination countries by origin
p <- ggplot(data = erasmus_tidier,
         aes(x = x, id = id, split = y, value = freq)) +
  geom_parallel_sets(aes(fill = sending_country_code),
                     alpha = 0.6, color = NA) +
  geom_parallel_sets_axes(axis.width = 0.1) +
  geom_parallel_sets_labels(color = NA, angle = 0, hjust = 0) +
  scale_fill_manual(values = countries_pal) +
  theme_void() +
  labs(subtitle = 'Receiving countries for students from the 5 countries\nthat send the most students on Erasmus visits') +
  theme(text = element_text(family = 'Lato'), 
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = 'none')

labeldata <- ggplot_build(p)$data[[3]]  # grab data for the third layer (parallel_sets_labels)

# add company labels
p +  # add origin country labels
  geom_text(data = labeldata %>%
              filter(PANEL == 1, x == 1),
              aes(x, y, label = label),
              inherit.aes = FALSE,
              nudge_x = -0.1,
              family = 'Lato', size = 3) +
  geom_text(data = labeldata %>%
              filter(PANEL == 1, x == 2),
              aes(x, y, label = label),
              inherit.aes = FALSE,
              nudge_x = 0.1,
              family = 'Lato', size = 3)

```

Next, I examined the 5 most commonly visited countries and examined the distribution of which countries those students visited from from.

```{r fig_sending, fig.height = 12, fig.width = 10}

# determine 5 countries with most visitors
top_countries <- erasmus %>%
  # only include international exchanges
  filter(sending_country_code != receiving_country_code) %>%
  group_by(receiving_country_code) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  slice_head(n=5)

# create a custom color palette based on the top 5 countries
countries_pal_2 <- c('FR' = '#002395', 
                   'BE' = '#ed2939',
                   'ES' = '#ffc400', 
                   'DE' = '#000000',
                   'IT' = '#009246')

# tally counts of destinations by origin country
# format data for ggforce
erasmus_tidier <- erasmus %>%
  # only include international exchanges
  filter(sending_country_code != receiving_country_code) %>%
  # only include top 5 countries as determined above
  filter(receiving_country_code %in% top_countries$receiving_country_code) %>%
  group_by(sending_country_code, receiving_country_code) %>%
  summarize(freq = n()) %>%
  ungroup() %>%
  gather_set_data(x = 1:2) %>%
  arrange(x, sending_country_code, receiving_country_code)
erasmus_tidier$x <- factor(erasmus_tidier$x,
                           levels = c('sending_country_code', 'receiving_country_code'))

# create parallel sets plot
# showing destination countries by origin
p <- ggplot(data = erasmus_tidier,
         aes(x = x, id = id, split = y, value = freq)) +
  geom_parallel_sets(aes(fill = sending_country_code),
                     alpha = 0.6, color = NA) +
  geom_parallel_sets_axes(axis.width = 0.1) +
  geom_parallel_sets_labels(color = NA, angle = 0, hjust = 0) +
  scale_fill_manual(values = countries_pal_2) +
  theme_void() +
  labs(subtitle = 'Sending countries for students who visit the 5 countries\nthat receive the most students on Erasmus visits') +
  theme(text = element_text(family = 'Lato'), 
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = 'none')

labeldata <- ggplot_build(p)$data[[3]]  # grab data for the third layer (parallel_sets_labels)

# add company labels
p +  # add origin country labels
  geom_text(data = labeldata %>%
              filter(PANEL == 1, x == 1),
              aes(x, y, label = label),
              inherit.aes = FALSE,
              nudge_x = -0.1,
              family = 'Lato', size = 3) +
  geom_text(data = labeldata %>%
              filter(PANEL == 1, x == 2),
              aes(x, y, label = label),
              inherit.aes = FALSE,
              nudge_x = 0.1,
              family = 'Lato', size = 3)

```


  